 \documentclass[11pt]{article}

 \usepackage{amsmath}
 \usepackage{amsxtra}
 \usepackage{amsfonts}
 \usepackage{amssymb}
 \usepackage{amsthm}
 \usepackage{graphicx}
 \usepackage{mathtools}
 %\usepackage{dsfont}
 \usepackage{hyperref}

 \graphicspath{ {/Users/andrew/hw/math191k/figures/}}

 \usepackage[margin=3cm]{geometry}
 \newcommand{\eps}{\varepsilon}
 \newcommand{\Z}{\mathbb{Z}}
 \newcommand{\N}{\mathbb{N}}
 \newcommand{\R}{\mathbb{R}}
 \newcommand{\Q}{\mathbb{Q}}
 \newcommand{\W}{\mathcal{W}}
 \newcommand{\U}{\mathcal{U}}
 \newcommand{\V}{\mathcal{V}}
 \newcommand{\C}{\mathbb{C}}
 \newcommand{\K}{\mathcal{K}}

 \newcommand{\set}[1]{\{ #1 \}}
 \newcommand{\gen}[1]{\left\langle #1 \right\rangle}
 \newcommand{\floor}[1]{\left\lfloor #1 \right\rfloor}
 \newcommand{\abs}[1]{\left| #1 \right|}
 \renewcommand{\phi}{\varphi}
 \renewcommand{\Re}{\operatorname{Re}}
 \renewcommand{\Im}{\operatorname{Im}}
 \newcommand{\conj}[1]{\mkern 1.5mu\overline{\mkern-1.5mu#1\mkern-1.5mu}\mkern 1.5mu}
 \newcommand{\defeq}{\vcentcolon=}
 \newcommand{\identity}{\mathds{1}}

 \DeclareMathOperator{\im}{im}
 \DeclareMathOperator{\res}{Res}
 \DeclareMathOperator{\Arg}{Arg}
 \DeclareMathOperator{\parg}{arg}
 \DeclareMathOperator{\Log}{Log}

 \theoremstyle{plain}
 \newtheorem{thm}{Theorem}

 \theoremstyle{definition}
 \newtheorem{recall}{Recall}
 \newtheorem{remark}{Remark}
 \newtheorem{definition}{Definition}
 \newtheorem{prop}{Proposition}
 \newtheorem{cor}{Corollary}
 \newtheorem{lemma}{Lemma}
 \newtheorem{ex}{Example}
 \newtheorem{claim}{Claim}
 \newtheorem{exercise}{Exercise}
 \newtheorem{notation}{Notation}
 \newtheorem{question}{Question}


 \title{Dynamical systems theory $\bigcap$ Knot theory}

 \author{Andrew Berger\\ Yitz Deng}


\begin{document}
\maketitle

\tableofcontents

\section{Crash course in dynamical systems theory}
This section may not be necessary/ may be able to reduce to a stub.

\subsection{Qualitative dynamics}

Attractors: stable/unstable fixed points, limit cycles, chaos


\section{Knots on Manifolds}

Fuck this is a whole thing - maybe restrict to dynamical systems on $\R^n$?


\begin{definition}[Link of closed orbits]
  Given a flow $\phi_t$ on a three-manifold $M$, the \emph{link of closed orbits} $L_\phi$ is defined to be the set of closed (periodic) oribts of $\phi_t$, considered as a (perhaps infinite) link in $M$
\end{definition}

Basic questions about $L_\phi$: (taken from Ghrist)

\section{Meaures of complexity}

Is it true that complicated dynamics give to rise to complicated knots, and simple dynamics give rise to simple knots? Sounds plausible, and is true with the proper formulation.
The measure of complexity for dynamical system we will use is topological entropy. Topological entropy essentially measures the growth rate of periodic orbits under increasing resolution.

\begin{definition}[Zero-entropy knots]
  The class of all knots generated from unknots via the operations of cabling and connected sums.
\end{definition}


\begin{thm}[Franks and Williams, 1985]
  If a $C^2$ smooth flow on $S^3$ (or $\R^3$) has positive topological entropy, then the link of periodic oribts is infinite and contains an infinite number of distinct knot types as components
\end{thm}

Ghrist remarks that it follows from the proof that any such flow must contain infintely many knots which are not in the class of zero-entropy knots, and that there must be knotted orbits of arbitrarily long length. An interesting
question is what is the behavior of such `asymptotically wild' knots on a compact manifold?



\subsection{Questions from this section}
\begin{itemize}
\item How `small' or `large' can this link be?
\item Does the existence of one type of knot force the coexistence of another?
\item \emph{Really} stupid question: can a flow with no limit cycles have a non-empty link of closed orbits? Definitionally it seems the answer is no, but Ghrist asks if its possible to create a finite link as the set of
closed oribts of a flow with no fixed points, which is confusing.
\item Does every chaotic system have positive topolgical entropy? Converse? Is chaotic system even well-defined?
\end{itemize}


\section{General Questions}

\begin{itemize}
    \item Can non-chaotic dynamical systems have knotted orbits? Yes, you can trivially construct a dynamical system having any knot $K$ as a limit cycle.
    \item Follow up question: can non-chaotic dynamical systems be universal? I doubt it
    \item Classification theorems for universality? That is, when does a dynamical system contain all knots as orbits.
    \item What is meant by `perhaps infinite' link in the definition of link of closed orbits? Wild link?
    \item Why does Ghrist seem to mostly only consider flows on $S^3$? Maybe because we are considering the periodic solutions to third order systems (which have 3-D solution space)
\end{itemize}

%%start here

\section{Introduction and background}

Initial exposition on the connection between dynamical systems theory and knot theory. The set of links.

\subsection{Templates}

Motivate: collapse out along stable manifolds


\begin{definition}
  Given a template $\tau$, we define
  \begin{itemize}
    \item Branch lines  $\set{l_j : j = 1, \cdots M}$
    \item Strips $\set{x_i : i = 1, \cdots N \geq 2 M}$
    \item Itinerary space $\Sigma_\tau = \set{a_0 a_1 a_2 \cdots} \subset \set{x_1, x_2, \ldots, x_N}^{\Z^+}$
    \item Transition matrix

      \begin{equation}
        A_\tau(i,j) = \left.\{ \
        \begin{smallmatrix}
          0: \not\exists \text{ a strip from } x_i \text{ to } x_j \\
          1: \exists \text{ a strip from } x_i \text{ to } x_j
        \end{smallmatrix}
      \end{equation}
  \end{itemize}
\end{definition}

Note that powers of $A_\tau$ capture allowable trajectories. (Can we expand this into saying anything interesting?)

Put short exposition on ordering $\vartriangleright$ of orbits


\subsection{Template renormalization}
Template renormalization exposition goes here. This should be done very carefully and in much greater detail than given in Ghrist, I think

\subsection{Mexico}
(What happens when you leave your computer unattended\dots)
mad max
hardcore henry
klean kanteen katie
burrito berger

\subsection{Braids}

\subsection{Link of knots}

\subsection{Renormalization}

\section{Outline of proof that $\mathcal{V}$ is a universal template}

\subsection{Braids and the theorem of Alexander}

Every link is isotopic to some closed braid on $P$ strands for some $P$. Cite the proper paper

\subsection{The templates $\W_q$}

An isotopic copy of any closed braid exists as a set of periodic oribts on some $W_q$ for sufficently large $q$.

\subsection{Construct $\W_{q+1}$ from $\W_q$}

Start with renormalized $\W_1 \in \V$, and append a pair of ears to\dots

\subsection{Find $\W_q \subset \V$ for all $q$}

This is very tricky and will have to be very loosely sketched




\section{Miscelaneous discussions and interesting open questions}




\end{document}
